# Introduction

```{r setup-intro}
source("scripts/setup.R")
```

Development of offshore wind energy will be essential for providing renewable energy to mitigate climate change impacts. Analysis of habitats and species in call and lease areas can allow for responsible siting that minimizes impacts on sensitive species and habitats and present opportunities for sensitive habitat protections within lease areas. Acoustic impacts from construction and operation may negatively impact some species [@mooneyACOUSTICIMPACTSOFFSHORE2022]. Hard substrate provided by the platforms may actually benefit other species [@wilsonHabitatcreationPotentialOffshore2009]. Although only 42 megawatts (MW) of offshore wind energy capacity is operational (in Rhode Island and New York) as of May 31, 2022, the pipeline of potential generating capacity is estimated to be 40 gigawatts (GW) -- enough to power 13 million homes [@musialOffshoreWindMarket2022].

Wind energy areas are under active development by the Bureau of Ocean Energy Management (BOEM) throughout the U.S. This report assesses habitats for sensitivity throughout the federal waters of the contiguous U.S. lower 48 states (excluding Alaska and Hawaii).

A couple of fundamental questions driving this analysis include:

1. _How different are zones (and regions)?_ (_**inter-zone/region**_)\
  Areas not already slated for the offshore wind leasing impact species and habitats less, so could be candidates for alternative future planning. Areas with designated military use and outside BOEM’s authority (e.g., National Marine Sanctuaries) need to be further excluded from consideration.

1.  _How different are blocks from each other within a given region?_ (_**inter-block**_, _**intra-zone**_)\
  Within the leasing process, the most oversight can be exercised within existing lease areas, especially ones in the earlier stages of Proposed Sale Notices and Call Areas. Blocks within the same zone highlight where the most conservation value exists relative to each other.

While density surfaces are most preferable, especially for determining number of individuals impacted by human activity (e.g., for Potential Biological Removal as mandated by the Marine Mammal Protection Act [@romanMarineMammalProtection2013]), the distributions of relatively few taxa have been described by these model types. The most comprehensive atlas of marine species distributions comes from the AquaMaps [@kaschnerAquaMapsPredictedRange2019], which uses only 4 terms per environmental predictor (min, max and preferred min and max) to produce a gradient of habitat suitability [@readyPredictingDistributionsMarine2010]. Yet another type of distributional information is based on a binary range map, such as produced by taxonomic experts with the IUCN RedList process. 

Collating the best available data has been a significant challenge to date and has become all the more important with the global push for 30% conservation protection of land and sea by 2030 with the Kunming-Montreal Global Biodiversity Framework resolution passed at the Convention of Biological Diversity meeting in Montreal in December 2022 [@findlayCOP15Crunch2023]. Although the U.S. was notably absent from the 190 signatory countries, President Biden has advocated for similar protections nationally through the "America the Beautiful" initiative. The lack of comprehensive biodiversity information has been noted [@carrollUSConservationAtlas2022].

<!-- 

TODO:

- Essential Fish Habitat (EFH) is too coarse, not preferred by [@friedlandResourceOccurrenceProductivity2021] (see [methods notes | offhab - Google Docs](https://docs.google.com/document/d/1JSJMBqqL4c44c8fx01I-elx7gK8dNyuABjLinaxC95k/edit#bookmark=id.sky4lrftqqq8)) 

- review [oharaAligningMarineSpecies2017]: compares AquaMaps with IUCN and brings up how aligns and misaligns. Points for Intro + Discussion.
-->


```{r}
dbDisconnect(con, shutdown = T)
```

